---
title: "Assignment 5"
output: html_notebook
---
1. The data set asthma comes from a study of the relationship between childhood asthma and exposure to
air pollution from concentrated animal feeding operations (CAFOs). For this problem, treat asthma
(Yes/No) as following a binomial distribution given exposure, and use a smooth version of logistic
regression for (a)-(c) below.

(a) Choose a criterion by which to select the smoothing parameter. Plot this criterion versus the
smoothing parameter and choose the optimal value for use in (b) and (c).


```{r}
df1 = data.frame(asthma)
library(stringr)
df1$Asthma = str_replace(df1$Asthma,"Yes",'1')
df1$Asthma = str_replace(df1$Asthma,"No",'0')
df1<-df1[order(df1$Exposure),]
attach(df1)
A<-as.numeric(as.character(df1$Asthma))
E<-df1$Exposure
fitl$deviance
library(locfit)
n <- length(A)
k <- seq(10,n,by=10)    #h = k/n 
cv <- mcr <- gcv <- rep(0,length(k))
for (i in 1:length(k)){
  fit <- locfit(A~lp(E,nn=k[i]/n,deg=1),family="binomial",ev=dat(cv=TRUE))
  pihat <- fitted(fit)
  yhat <- (pihat > .5)*1              
  mcr[i] <- mean(abs(A-yhat))      
  cv[i] <- (-2*(sum(log(pihat[A==1 ]))+sum(log(1-pihat[A==0 ])))/n)   
  #  gcv[i] <- gcv(fit)["gcv"]
  gcv[i]<- (-2*n*fit$dp["lk"]/(n-fit$dp["df2"])^2)
}
par(mfrow=c(1,2))
plot(k,mcr,type="l",xlab = 'index',ylab="Misclassification rate", ylim=c(.0,.2))
plot(k,cv,type="l",xlab = 'index',ylab="CV using deviance")

par(mfrow=c(1,1))
plot(k,cv,type="l",xlab = 'index',ylab="CV using deviance",main="Cv vs GCV",ylim=c(0.0,1.5))
lines(k,gcv,type="l",xlab = 'index',ylab="GCV", col='blue')
```

(b) Plot a smooth curve estimating the relationship between exposure and the log-odds of developing asthma.


```{r}
h_cv<-k[which.min(cv)]
which.min(cv)
h_gcv<-k[which.min(gcv)]
which.min(gcv)
fit <-locfit(A~lp(E,nn=h_gcv/n,deg = 1),family="binomial",ev=dat(cv=TRUE))
pihat <- fitted(fit)
plot(E,log(pihat/(1-pihat)),type = "l",xlab = 'Exposure',ylab ='log-odds of developing asthma')

```
(c) Plot a smooth curve estimating the relationship between exposure and the probability of developing
asthma, with confidence bands.


```{r}
plot(E,fitted(fit),type="l",xlab='Exposure',ylab='probability of developing asthma')
plot(fit,band='local',get.data = T,main='with degree=1',xlab = 'Exposure',ylab='Asthma')

```

(d) Prepare an ANOVA table (or rather, an analysis of deviance table) testing the sequence of models:
Null ⊂ Linear ⊂ Nonlinear

```{r}
fit1<-lm(A~1,family="binomial")
fit2<-lm(A~E,family="binomial")
fit3 <- ksmooth(A,E,kernel="normal",bandwidth=30)
anova(lm(fit3$y~fit3$x))
anova(fit1,fit2)
```

3. The library ISLR contains a dataset called Wage with information about wages and demographic factors
( age, gender, health, education, etc). Fit three splines to estimate the relationship between age and
wage: natural base, B-spline and smooth-splines.


```{r}
library(ISLR)
names(Wage)
str(Wage)
Wage<-Wage[order(Wage$age),]
head(Wage)
attach(Wage)


```
(a) State the criterion by which to select the smoothing parameter and use it to select. Plot this
criterion versus the smoothing parameter and choose the optimal value for use in (b) and (c).
```{r}
#Plotting the Regression Line to the scatterplot  
library(splines)
fit<-lm(Wage$wage~bs(Wage$age))
plot(Wage$age,Wage$wage,col="grey",xlab="Age",ylab="Wages",main = "Regression plot")
lines(Wage$age,predict(fit),col="darkgreen",lwd=2,type="l")

lm.fit=lm(wage~age,data=Wage)
coef(lm.fit)

plot(age,wage,col = 'grey')
lines(smooth.spline(age,wage,df=10),lwd=2,col="blue")
lines(smooth.spline(age,wage,df=61),lwd=2,col="green")


```


(b) Plot a smooth curve estimating the relationship between age and the wages.

```{r}
plot(age,wage,col='grey')
spline_fit_10=smooth.spline(Wage$age,Wage$wage,df=10)
spline_fit_CV=smooth.spline(Wage$age,Wage$wage,cv=TRUE)
spline_fit_CV$df
lines(spline_fit_10$x,spline_fit_10$y,lwd=2,col='red')
lines(spline_fit_CV$x,spline_fit_CV$y,lwd=2,col='blue')

```

(c) Compare the three models.

```{r}
plot(age,wage,col='gray')
#Natural spline model 
fit1<-lm(wage~ns(age,df=10))
lines(age,predict(fit1),lwd =2,col='yellow')
#B-spline model
fit2<-lm(wage~bs(age,df=10))
lines(age,predict(fit2),lwd =2,col='blue')
#Smoothing spline model
#plot(age,wage,col='gray')
lines(smooth.spline(age,wage,df=10),lwd =2,col='Red')
lines(smooth.spline(age,wage,df=61),lwd =2,col='black')
legend("topright",c("Natural Spline","B-spline","Smoothing Spline with 10 df","Smoothing Spline with 61 df"),col=c("yellow","blue","red","black"),lwd=2)
```

(d) Use B-splines to create 4 plots corresponding to splines with degrees of freedom = p × number of
points, for the following p’s: p = 0.05, 0.5, 0.75, 0.98.

```{r}
n=length(age)
df1=floor(.05*n)
#plot(age,wage,col="grey")
fit3<-lm(wage~bs(age,df=df1))
plot(age,wage,col="grey",main=paste('df=',df1,' (n points)'))
lines(age,predict(fit3),lwd =2,col='blue')

df2 = floor(.5*n)
fit4<-lm(wage~bs(age,df=df2))
plot(age,wage,col="grey",main=paste('df=',df2,' (n points)'))
lines(age,predict(fit4),lwd =2,col='green')

df3 = floor(.75*n)
fit5<-lm(wage~bs(age,df=df3))
plot(age,wage,col="grey",main=paste('df=',df3,' (n points)'))
lines(age,predict(fit5),lwd =2,col='yellow')

df4=floor(.98*n)
fit6<-lm(wage~bs(age,df=df4))
plot(age,wage,col="grey",main=paste('df=',df4,' (n points)'))
lines(age,predict(fit6),lwd =2,col='red')
```
4 (a) Write the set of truncated spline basis functions for representing a cubic spline function with three knots inside [0, 1].

```{r}
##
library(splines)
X <- seq(0, 1, length=100)
knots <- c(0.25, 0.50, 0.75)
n=length(x)
set.seed(1); err=1
y<- (2.3*cos(3.0*x) + 1.2*sin( 4.5*x) + cos(1.92*x)  + rnorm(n, 0, err))


```

(b) For x ∈ [0, 1] and equally spaced knots, plot the above basis functions (exclude the intercept term).
```{r}
plot(x,y, col="grey")
lines(x, predict(lm(y~ns(x, df=10,knots=c(0.25, 0.50, 0.75),intercept=FALSE))), col='blue')
```

(c) For the same range and knots as in (b), plot the B–spline basis functions, again excluding the
intercept term. You may use the splines package to construct them for you.
```{r}
library(splines)
spl <- bs(x,df=10,knots=c(0.25, 0.50, 0.75),intercept=FALSE)  
plot(spl[,1]~x, ylim=c(0,max(spl)), type='l', lwd=2, col=1, 
     xlab="B-spline basis", ylab="")
for (j in 2:ncol(spl)) lines(spl[,j]~x, lwd=2, col=j)
```


